---
title: "P04 AC8 OE membranes +/- Calmodulin - TRYPTIC CONTROLS"
author: "Dina"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: spacelab
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: hide
---
<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Experiment

Experiment performed on: **220929**
 
Membrane suspensions were prepared on 220913, snap-frozen and stored in liquid nitrogen.
Buffer composition: 100 mM HEPES-KOH pH 7.4, 150 mM KCl, 1 mM MgCl2( + 1 mM CaCl2 added before experiment)

CaM stock: 4 mg/mL in 50 mM HEPES-NaOH pH 7.5, 150 mM NaCl (add CaCl2 to 1 mM)

Dilute CaM with 50 mM HEPES-NaOH pH 7.5, 150 mM NaCl, 1 mM CaCl2

Blank: 50 mM HEPES-NaOH pH 7.5, 150 mM NaCl, 1 mM CaCl2

CaM mounts spiked in: 

* 0 ug
* 3 ug

All experiments were conducted in quadruplicates. 
The suspensions were incubated with CaM for 10 minutes at 25°C, followed by a treatment with water (TC). The sample was then heated to 99°C for 5 minutes, cooled down at 4°C for 5 minutes, then treated with the same volume of 10 % sodium deoxycholate (standard LiP procedure). 
The LiP step was followed by a tryptic digest over night and a C18 clean-up the next day. 
A library of pooled replicates was produced. 
After drying, peptides were resuspended and measured on a **Thermo Eploris 480** mass spectrometer on **220930**.
Data was searched in Spectronaut (v.15).
A contaminant FASTA file was included, single hits were excluded, 5 AA were set as the minimum peptide length. Data was median normalized in Spectronaut. Imputation was disabled. Peptides are grouped by their modified sequences.


# PREPARATION

## Loading packages

**protti** is an R package for LiP-MS and bottom-up proteomics data analysis developed in the Picotti lab (see here: [GitHub](https://github.com/jpquast/protti), [publication](https://doi.org/10.1093/bioadv/vbab041)).

Other packages used are **tidyverse** and **data.table**.


```{r loading_packages, message=FALSE, warning=FALSE, error=FALSE}
library(protti) # v.0.5.0
library(tidyverse) # v.1.3.1
library(data.table) # v.1.14.2

setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # set working directory to location of R script
```

## Loading data

```{r loading_data, message=FALSE, warning=FALSE}
DIA_raw <- read_protti("20221007_142430_dschuster CNG TC_Report.csv")

```
Only proteotypic peptides will be retained, except for CNG proteins and Calmodulin.

```{r cleaning_data, message = FALSE, warning = FALSE}

DIA_raw_filtered <- DIA_raw %>%
  filter(pg_fasta_files != "contaminants", pg_fasta_files != "NaN") %>%
  filter(fg_quantity > 10) %>%
  mutate(intensity_log2 = log2(pg_quantity))

DIA_clean <- DIA_raw_filtered %>% 
  separate(r_condition, c("digestion", "concentration"), " ", remove = FALSE) %>%
  mutate(concentration = as.numeric(concentration))  %>%
  mutate(pg_protein_accessions = ifelse(pg_protein_accessions %like% "Q00194", "Q00194", pg_protein_accessions)) %>%
  mutate(pg_protein_accessions = ifelse(pg_protein_accessions %like% "Q28181", "Q28181", pg_protein_accessions)) %>%
  mutate(pg_protein_accessions = ifelse(pg_protein_accessions %like% "P62157", "P62157", pg_protein_accessions))

DIA_clean_proteotypic <- DIA_clean %>%
  filter(pg_protein_accessions!= "Q00194" | pg_protein_accessions!= "Q28181" |  pg_protein_accessions!= "P62157") %>%
  filter(pep_is_proteotypic == T)

DIA_clean_CNG_non_proteotypic  <- DIA_clean %>%
  filter(pg_protein_accessions %like% "Q00194" | pg_protein_accessions %like% "Q28181" | pg_protein_accessions %like% "P62157")

DIA_clean_final <- DIA_clean_proteotypic %>%
  rbind(DIA_clean_CNG_non_proteotypic) %>%
  mutate(uniprot_id = pg_protein_accessions) %>%
  select(digestion, 
         concentration,
         r_condition,
         r_file_name, 
         r_replicate, 
         pg_fasta_files, 
         uniprot_id, 
         pep_ms2quantity,
         intensity_log2,
         pep_grouping_key,
         pep_stripped_sequence,
         pep_nr_of_missed_cleavages
         )
  
unis <- unique(DIA_clean_final$uniprot_id)

uniprot <-
  fetch_uniprot(
    unis,
    columns =  c(
      "protein_name",
      "gene_names",
      "length",
      "sequence"
    )
  ) %>%
  rename(
    protein_sequence = sequence,
    length_protein = length,
    uniprot_id = accession
  )

DIA_clean_uniprot <- DIA_clean_final %>%
  left_join(uniprot, by = "uniprot_id") %>%
  find_peptide(protein_sequence, pep_stripped_sequence) %>%
  peptide_type(aa_before, last_aa) %>%
  sequence_coverage(protein_sequence, pep_stripped_sequence) %>%
  distinct()

proteins_identified <- uniprot %>%
  distinct(uniprot_id)

DIA_clean_CNG <- DIA_clean_uniprot %>%
  filter(uniprot_id == "Q00194" | uniprot_id == "Q28181")

```


# DATA ANALYSIS


## Volcano plots

Volcano plots are shown to compare the protein abundances between the untreated and CaM treated condition using the [proDA](https://www.biorxiv.org/content/10.1101/661496v2) algorithm.
p-values are calculated adjusted for multiple testing with Benjamini-Hochberg correction.
An FDR cutoff of 0.01 and a log2(fold change) cutoff of 1 is applied.
CNG proteins (Q00194, Q28181) and Calmodulin (P62157) are highlighted in the plot below.


```{r Volcano_plot_input, message = FALSE, warning = FALSE}
Volcano_input <- DIA_clean_uniprot %>%
  dplyr::select(
    uniprot_id,
    r_file_name,
    r_condition,
    r_replicate,
    protein_name,
    intensity_log2,
    start,
    end,
    pep_type,
    protein_sequence,
    length_protein
  ) %>%
  filter(r_condition == "TC 0" | r_condition == "TC 3")

Volcano_input <- Volcano_input %>%
  unique() %>%
  assign_missingness(r_file_name,
                     r_condition,
                     uniprot_id,
                     intensity_log2,
                     ref_condition = "TC 0",
                     retain_columns = c(protein_sequence, length_protein))

t_test_result_proDA_prot <- diff_abundance(Volcano_input, 
                                r_file_name, 
                                r_condition, 
                                uniprot_id,
                                intensity_log2,
                                missingness, 
                                comparison, 
                                ref_condition = "TC 0", 
                                method = "proDA", 
                                retain_columns = c(protein_sequence, length_protein))

```


```{r volcano_plot_AC8, fig.align= "center", message = FALSE, warning = FALSE}

volcano_plot(t_test_result_proDA_prot, uniprot_id, diff, adj_pval, method = "target", target_column = uniprot_id, target = c("P62157", "Q00194", "Q28181"),  x_axis_label = "log2(fold change) 3 ug CaM vs 0 ug CaM", title = "Volcano plot protein level", y_axis_label = "-log10(q-value)", log2FC_cutoff = 1, significance_cutoff = 0.01, interactive = TRUE)

```

# SUPPLEMENT - Quality control 

## Tryptic control

### Coefficients of variation

```{r quality_control_cvs, fig.width = 12, fig.align= "center",  message = FALSE, warning = FALSE}
qc_cvs(
  data = DIA_clean_uniprot,
  grouping = pep_grouping_key,
  condition = r_condition,
  intensity = pep_ms2quantity,
  plot = FALSE
)

qc_cvs(
  data = DIA_clean_uniprot,
  grouping = pep_grouping_key,
  condition = r_condition,
  intensity = pep_ms2quantity,
  plot_style = "density",
  plot = TRUE
)

qc_cvs(
  data = DIA_clean_uniprot,
  grouping = pep_grouping_key,
  condition = r_condition,
  intensity = pep_ms2quantity,
  plot_style = "violin",
  plot = TRUE
)

```

### Intensity distribution

```{r quality_control_log2_int_distr, fig.align= "center", message = FALSE, warning = FALSE}
qc_intensity_distribution(
  DIA_clean_uniprot,
  r_file_name,
  pep_grouping_key,
  intensity_log2,
  plot_style = "histogram"
)

```

### Missed cleavages

```{r quality_control_missed_cleav, fig.align= "center", message = FALSE, warning = FALSE}
qc_missed_cleavages(
  DIA_clean_uniprot,
  r_file_name,
  pep_grouping_key,
  pep_nr_of_missed_cleavages,
  pep_ms2quantity,
  method = "intensity",
  plot = TRUE,
  interactive = FALSE
)

qc_missed_cleavages(
  DIA_clean_uniprot,
  r_file_name,
  pep_grouping_key,
  pep_nr_of_missed_cleavages,
  pep_ms2quantity,
  method = "count",
  plot = TRUE,
  interactive = FALSE

)
```

### Peptide types

```{r quality_control_pep_type, fig.align = "center", message = FALSE, warning = FALSE}
qc_peptide_type(
  DIA_clean_uniprot,
  r_file_name,
  pep_grouping_key,
  pep_type,
  intensity = pep_ms2quantity,
  method = "count",
  plot = TRUE,
  interactive = FALSE
)

qc_peptide_type(
  DIA_clean_uniprot,
  r_file_name,
  pep_grouping_key,
  pep_type,
  intensity = pep_ms2quantity,
  method = "intensity",
  plot = TRUE,
  interactive = FALSE
)

```

### Number of protein IDs per sample

```{r quality_control_pep_ID, fig.align = "center", message = FALSE, warning = FALSE}
qc_ids(DIA_clean_uniprot, r_file_name, uniprot_id, condition = r_condition, intensity = intensity_log2)

```

### Principal component analysis (PCA)

```{r PCA, message = FALSE, warning = FALSE}
qc_pca(DIA_clean_uniprot, 
       r_file_name, 
       uniprot_id, 
       intensity_log2, 
       r_condition)

```
